Rising from the Ashes II: The Bar-driven Abundance Bimodality of the Milky Way

### 2025-02-20 paper/ms.tex
Bow down to AI overlords
Add thick disk papers

The bimodality inherits a long history of attempts to decompose the disk dating back to \citet{1983MNRAS.202.1025G}, who showed that the vertical distribution of stars is well-fit by a double exponential. This led to a separation based on kinematics between the ``thin'' and ``thick'' disk. It was later shown that the thick disk is more $\alpha$-enhanced than the thin disk \citep{1996ASPC...92..307G,1998A&A...338..161F,2004AN....325....3F,2006MNRAS.367.1329R}. However, it was not until more recently that it became obvious there is a clean separation between two sequences in the abundance plane, i.e., the $\alpha$-rich and $\alpha$-door disk \citep{2011A&A...535L..11A,2012A&A...545A..32A,2014A&A...562A..71B,2014ApJ...796...38N,2020MNRAS.493.2952H}, also shown in the upper left panel of Figure~\ref{fig:fig1}. The Milky Way is the only galaxy for which such a structure has been definitively shown to exist.\footnote{Note claims of a detection \citep{2023ApJ...956L..14K} and nondetection \citep{2024IAUS..377..115N} in M31.}

@@

>>>>

\edit1{The bimodality inherits a long history of attempts to decompose the disk dating back to \citet{1983MNRAS.202.1025G}, who showed that the vertical distribution of stars is well-fit by a double exponential. This led to a separation based on kinematics between the ``thin'' and ``thick'' disk. It was later shown that the thick disk is more $\alpha$-enhanced than the thin disk \citep{1996ASPC...92..307G,1998A&A...338..161F,2004AN....325....3F,2006MNRAS.367.1329R}. However, it was not until more recently that it became obvious there is a clean separation between two sequences in the abundance plane, i.e., the $\alpha$-rich and $\alpha$-door disk \citep{2011A&A...535L..11A,2012A&A...545A..32A,2014A&A...562A..71B,2014ApJ...796...38N,2020MNRAS.493.2952H}, also shown in the upper left panel of Figure~\ref{fig:fig1}. The Milky Way is the only galaxy for which such a structure has been definitively shown to exist.\footnote{Note claims of a detection \citep{2023ApJ...956L..14K} and nondetection \citep{2024IAUS..377..115N} in M31.}}

\edit1{The distinction between a chemical and kinematic decomposition of the disk is illustrated by a population of low-$\alpha$ stars on vertically extended orbits in the outer disk \citep{2015ApJ...808..132H,2016ApJ...823...30B}. This is thought to arise from flaring associated with radial migration \citep{2015ApJ...804L...9M,2016ApJ...831..139M}, or from other processes, including misaligned gas accretion and minor mergers \citep{2010MNRAS.408..783R,2009MNRAS.396..696S}. Nonetheless, in the outer disk the kinematically-defined thick disk has contributions from both the chemically defined high- and low-$\alpha$ sequences.}

@@

\edit1{We acknowledge the use of OpenAI’s ChatGPT and Anthropic's Claude for assistance in editing this manuscript for clarity and conciseness and in generating small analysis scripts and code snippets.}
